---
title: "Iteration with loops"
author: "Le Nguyen"
date: "Module 4 A2 Week 4"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

<style type="text/css">
/* Title */
h1.title {
  color: #ED4F4F;
  font-size:35px;
  font-weight: bold;
}
/* Level 1 header */
h1 {
  color: #4053B7;
}
/* Level 2 header */
h2 {
  color: #4053B7;
}
/* Table of contents */
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #4053B7;
    border-color: #337ab7;
}
/* Navigation pills */
.nav-pills>li.active>a, .nav-pills>li.active>a:focus, .nav-pills>li.active>a:hover {
    color: #fff;
    background-color: #ED4F4F;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA)
```

# Function `fib()`

## Introduction

The Fibonacci sequence is the sequence $\{F_n\colon n \geq 1\}$ of numbers
defined by $F_1 = 0, F_2 = 1$ and $F_n = F_{n-2} + F_{n-1}$ for $n\geq 3$. 
For example, the first seven Fibonacci numbers are
\begin{align*}
F_1 &= 0\\
F_2 &= 1\\
F_3 &= 0+1 = 1\\
F_4 &= 1 + 1 = 2\\
F_5 &= 1 + 2 = 3\\
F_6 &= 2 + 3 = 5\\
F_7 &= 3 + 5 = 8\\
\end{align*}

## Part a

### `fib()`

Write a function, `fib()`, which uses a `for` loop. Function `fib()` should 
take one argument, `n`, and return the first `n` Fibonacci numbers in a vector.
See below for some examples.

```{r fib-function}
fib = function(n){
  
  num1 = 0
  num2 = 1
  num3 = 0
  fib = c(num1,num2)
  for (i in 1:n) {
    
    num3 = num1 + num2
    fib = c(fib,num3)
    num1 = num2
    num2 = num3
  }
  
  return(fib)
}

fib(5)


```

```{r fib-function-test, eval=FALSE, results='hold'}
fib(n = 1)
fib(n = 2)
fib(n = 7)
fib(n = 10)
```

## Part b

### `fib_1()`

Modify `fib()` so that it uses a `while` loop instead of a `for` loop. Call
this new function `fib_1()`. Test that your function works with the following
examples below.

```{r fib-function-while}
fib_1 = function(n){
  
  num1 = 0
  num2 = 1
  num3 = 0
  fib = c(num1,num2)
  i = 0
  while( i != n) {
    
    num3 = num1 + num2
    fib = c(fib,num3)
    num1 = num2
    num2 = num3
    i = i +1
  }
  
  return(fib)
}

fib_1(5)


```

```{r fib-function-test-while, eval=FALSE, results='hold'}
fib_1(n = 1)
fib_1(n = 2)
fib_1(n = 7)
fib_1(n = 10)
```


# Function `all.primes()`

## Introduction

Your goal is to write a function `all.primes()` that will return all possible
prime numbers among the first `n` positive integers. For example, among the 
integers 1 through 10, there are prime numbers: 2, 3, 5, 7. Recall that
a prime number is a whole number greater than 1 whose only factors are 
1 and itself.

To get started, proceed to step 1.

## Step 1

### `is.prime()`

R has built-in functions that perform logical tests. For example, 
`is.numeric()` and `is.character()` performs a logical test if an object is
numeric and if an object is of type character, respectively. Write a function
called `is.prime()` that tests if a numeric vector of length 1 is a 
prime number or not. The function should return a single `TRUE` or `FALSE`
value. See below for some examples.

```{r is.prime}

is.prime = function(x){
  for (i in 2:(x-1)) {
    test = x %% i
    if (x ==2 ){
      prime = TRUE
      break
    }else if (test == 0){
      prime = FALSE
      break
    }else{
      prime = TRUE
    }
  }
  return(prime)
}


```

```{r is.prime-examples, eval=FALSE, results='hold'}
is.prime(x = 2)
is.prime(x = 10)
is.prime(x = 13)
```


## Step 2 

### `all.primes()`

Write a function `all.primes()` that has one argument, `n`, and returns all
possible prime numbers among the first `n` positive integers. See below for
some examples.

```{r all.prime}
all.primes <- function(n) {
 x = seq(2, n)
 primes = c()
 for (i in seq(2, n)) {
 if (any(x == i)) {
 primes = c(primes, i)
 x = c(x[(x %% i) != 0], i)
 }
 }
 return(primes)
 }

all.primes(13)

#I actually got stuck on this problem problem, so I looked at a soltuion to it: https://www.w3resource.com/r-programming-exercises/basic/r-programming-basic-exercise-6.php

```

```{r allprimes-examples, eval=FALSE, results='hold'}
all.primes(n = 10)
all.primes(n = 21)
all.primes(n = 2)
```
