---
title: "Using joins to create MLB tables"
author: ""
date: "Module 7 A1 Week 7"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

<style type="text/css">
/* Title */
h1.title {
  color: #000088;
  font-size:60px;
  font-weight: bold;
}
/* Level 1 header */
h1 {
  color: #000088;
  font-weight: bold;
}
/* Level 2 header */
h2 {
  color: #CC0000;
  font-weight: bold;
}
/* Level 4 header */
h4 {
  color: #000000;
  font-weight: bold;
}
/* Table of contents */
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #CC0000;
    border-color: #337ab7;
}

</style> 

```{r setup, include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA,
                      fig.width=9, 
                      fig.height=7)
```

# Introduction

Today you will get practice merging data frames with inner and outer join
functions available in package `dplyr`. To get started, load packages
`tidyverse` and `Lahman`.

Package `Lahman` has numerous data frames about Major League Baseball.
Type `help(package = "Lahman")` in your Console to see everything available.

```{r packages}
library(tidyverse)
library(Lahman)
```

```{r hidden-packages, include=FALSE}
# library(extrafont)
# loadfonts(device = "win")
```

```{r theme-custom, include=FALSE}
theme_custom <- function() {
   theme_bw() +
   theme(#text = element_text(family = "Comic Sans MS", color = "#000088"),
         axis.title = element_text(size = 16), 
         title = element_text(size = 20),
         legend.title = element_text(size = 10),
         legend.text = element_text(size = 10),
         axis.text.x = element_text(size = 12),
         axis.text.y = element_text(size = 12),
         plot.caption = element_text(size = 10))
 }
```

# Data

You will work with data frames in package `Lahman`. When needed, utilize R's
help to get an understanding of what the variables are in a given data frame.
For example, `?Salaries` will provide a short description for each of the 
variables in data frame `Salaries`.

# Questions

## Question 1 

Select three data frames from package `Lahman`. Identify what variables are
in common between any pair of the three data frames, and identify what
variables are in common between all three data frames. What are the primary
keys for each data frame?

```{r}
LahmanData
names(AllstarFull)
names(AwardsPlayers)
names(Batting)

#Between AllstarFull and AwardsPlayers: playerID, yearID
#Between All: PlayerID, yearID
#Primary Key is PlayerID

```
Between AllstarFull and AwardsPlayers: playerID, yearID
Between All: PlayerID, yearID
Primary Key is PlayerID for all



## Question 2 

Use data frames `Managers` and `AwardsManagers` to reproduce the 
data frame given below.


```{r}
temp = Managers %>%
  filter(yearID == "2016") %>%
  right_join(AwardsManagers, by = "yearID")
temp
```


## Question 3

Use data frames `Managers`, `AwardsManagers`, and `Master` to reproduce the 
data frame given below.


```{r}
temp2 = Managers %>%
  left_join(AwardsManagers) %>%
  left_join(Master, by = "playerID") %>%
  filter(yearID >= "2000") %>%
  select('yearID','teamID','nameFirst','nameLast','awardID') 

temp2
```


## Question 4

Use data frames `Teams` and `Salaries` to reproduce the plot below. Bar colors
represent each team's primary color.

```{r}
winners = Teams %>%
  filter(LgWin == 'Y') %>%
  filter(yearID >= '2000') %>%
  filter(yearID <= '2016') %>%
  select(teamID,yearID)
   
win = semi_join(Salaries,winners)

ggplot(win, aes(x = teamID, y = salary/10e6)) + geom_bar(stat = 'identity') + coord_flip() + scale_x_discrete() + labs(y ="Salary (millions)") 

```


## Question 5

Adjust your plot in Question 4 for inflation with year 2000 as baseline.
Comment on the differences between the plots.

```{r}
year = seq(2000,2016,1)
inflation = c(1.00,1.03,1.04,1.07,1.10,1.13,1.17,1.20,1.25,1.25,1.27,1.31,1.33,1.35,1.37,1.38,1.39)
 
win2 = win

for (i in 1:16)
{
  for (j in 1:length(win2$yearID)){
  if (win2$yearID[j] == year[i]){
    win2$Salary = win2$salary/inflation[i]
  }
  
  }
}

ggplot(win2, aes(x = teamID, y = Salary/10e6)) + geom_bar(stat = 'identity') + coord_flip() + scale_x_discrete() + labs(y ="Salary (millions)") 


```
I don't know if my approach was correct, especially with my data processesing, but from what I can see, even after inflation, the porpotion of sallaries are still the same. 


Year | USD Value | Inflation Rate
-----|-----------|----------------
2000|	$1.00|	3.36%
2001|	$1.03|	2.85%
2002|	$1.04|	1.58%
2003|	$1.07|	2.28%
2004|	$1.10|	2.66%
2005|	$1.13|	3.39%
2006|	$1.17|	3.23%
2007|	$1.20|	2.85%
2008|	$1.25|	3.84%
2009|	$1.25|	-0.36%
2010|	$1.27|	1.64%
2011|	$1.31|	3.16%
2012|	$1.33|	2.07%
2013|	$1.35|	1.46%
2014|	$1.37|	1.62%
2015|	$1.38|	0.12%
2016|	$1.39|	1.26%

# References {#References-link}

1. Lahman, S. (2017) Lahman's Baseball Database, 1871-2016.

2. RStudio Cheatsheets