---
title: "Simple linear regression"
author: "Analyzing the price of a car"
date: ""
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

<style type="text/css">
/* Title */
h1.title {
  color: #206005;
  font-size:40px;
  font-weight: bold;
}
/* Level 1 header */
h1 {
  color: #366f1d;
}
/* Level 2 header */
h2 {
  color: #366f1d;
}
/* Level 4 header */
h4 {
  font-weight: bold;
}
/* Table of contents */
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #a5bf9b;
    border-color: #337ab7;
}
    
</style>

```{r setup, include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA,
                      fig.width=6, 
                      fig.height=5)
```

## Introduction

This group assignment will use the Cars93 dataset from the MASS package and is inspired from the Lab 8 written by CMU's Professor Alexandra Chouldechova for her Programming in R for Analytics course.

Packages to be used:

1. `MASS`
2. `tidyverse`
3. `broom`

```{r packages, echo=TRUE}
library(MASS)
library(tidyverse)
library(broom)
library(dplyr)
```

## Data

The Cars93 dataset consists of data from 93 cars on sale in the USA in 1993. It has 27 variables.


```{r data, echo=TRUE}
attach(MASS::Cars93)
Cars93 <- as_tibble(MASS::Cars93)
glimpse(Cars93)
levels(AirBags)
```

#### The variables

```{r glimpse.data, echo=FALSE}
glimpse(Cars93)
```

## Exploratory data analysis

#### Question 1

Create a scatter plot between `Price` and the `EngineSize`. Comment
on the relationship. What do you think the correlation is between the variables?

```{r Q1, fig.align='center', fig.height = 4, echo=FALSE}
#Code for Q1
ggplot(Cars93, aes(x = EngineSize, y = Price)) + geom_point()

#There appears to be a positive correlation between engine size and price
```

#### Question 2

Create scatter plots between `Price` and categorical variables `Airbags` and `Origin`. Comment on the relationship. Add `geom_jitter()` to your plot. What do you notice about the variability in `Price` between different levels?


```{r Q2, fig.align='center', fig.height = 4, echo=FALSE}
#Code for Q2

ggplot(Cars93, aes(x = AirBags, y = Price)) + geom_point() + geom_jitter()
ggplot(Cars93, aes(x = Origin, y = Price)) + geom_point() + geom_jitter()

#For airbags, the more airbags in the car the more variability in price. For Origin, non-usa cars have the most variability. 

```

## Model fit

#### Question 3

Fit a linear model with function lm() between variables `Price` and `EngineSize`. Save it as an object mod.fit. What are the values of $b_0$ and $b_1$? Write out the model. *Hint*: `tidy()`

```{r Q3, echo=FALSE}
#Code for Q3
Cars93 %>% 
  ggplot( aes(x = EngineSize, y = Price)) + geom_point() + geom_smooth(method=lm, se= FALSE)
mod.fit <- lm(EngineSize ~ Price, data = Cars93)
mod.fit %>%
  broom::tidy()
#b0 = 1.41, b1 = .064
```

#### Question 4

Interpret the slope of the fitted model. Does the intercept have any practical meaning within the scope of our data?
```{r}
#The slope of the fit indicates how the price changes with engine size. The intercept mathematically would represent the price if the engine size is 0, but that is meaningless in this context.  


```



#### Question 5 

Plot the residual plot and comment on the model fit. *Hint*: `augment()`

```{r Q5, echo=FALSE}
#code for Q5

mod.fit %>%
  augment() %>%
  ggplot(mapping = aes(x =EngineSize,y = .resid)) +geom_point()

#It looks like there is a trend in the residuals, they are definitely not random, so there is a better fit to the data than linear. 

```


#### Question 6 

How much of the variability in the outcome variable `Price` is explained by the regression equation? Comment on the $R^2$. *Hint*: `glance()`

```{r Q6, echo=FALSE}
#code for Q6

mod.fit %>%
  glance(r.squared) 

#The r^2 value here is low, so there is high variability in the linear fit which means it's not very good.
  
```

#### Question 7 

Compute the sum of squared residuals. Extract the residuals using `augment()`. Match these residuals to the original data and sort the residuals. Is there manufacturers that on averages produces larger or smaller residuals? What do you notice about the residual patterns with respect to the manufacturers?

```{r Q7, echo=FALSE}
#code for Q7
res = resid(mod.fit)
sum_res = sum(res**2)
sum_res
Cars93
plot(data$Manufacturer,res, las = 2)

#Yes, some manufacturers produce larger average residuals. The highest ones tend to be domestically made cars with a high range of residuals- meaning high range in price. 

```

## Exploring additional variables


Below is figure showing how Price varies with EngineSize in the Cars93, with accompanying regression lines. There are two plots, one for USA cars, and one for non-USA cars.

```{r, fig.align='center', fig.height = 4, echo=FALSE}
qplot(data = Cars93, x = EngineSize, y = Price, colour = Origin)+ 
  facet_wrap("Origin") + 
  stat_smooth(method = "lm") + 
  theme(legend.position="none")
```

#### Question 8

Use the `lm()` function and fit two models to regress Price on EngineSize with respect to the `Origin` (US and non-US).
*Hint*: `filter()`

```{r Q8, echo=FALSE}
# US origin
data = Cars93

data %>%
  filter(Origin == "USA") %>%
  ggplot( aes(x = EngineSize, y = Price)) +
  ggtitle("USA")+
  geom_point()+
  stat_smooth(method = "lm") 
  
# For non-US origin
data %>%
  filter(Origin == "non-USA") %>%
  ggplot( aes(x = EngineSize, y = Price)) +
  ggtitle("non-USA")+
  geom_point()+
  stat_smooth(method = "lm")
```


#### Question 9

Compare and contrast the fitted models in Question 8 (1-4 lines).
```{r}
#The price is positively correlated with engine size in both USA and non-USA cars. USA cars a larger spread in prices, while non-USA has a larger range in prices. 
```


#### Question 10

Multiple regression and categorical predictors were briefly introduced in the class. Do you think analyzing the price based on the size of the engine and the origin will give us more information than the model fitted in Question 3? Justify your answer.

```{r}
#I think separating by origin of the car will give us more information because the fits are more accurate. We know this from looks at the residuals and seeing that USA cars have a much larger range in residuals. This makes trying to fit to all the cars at once very inaccurate because the spread of the data is different depending on origin. With separating by category we get better fits, which gives us a better look at the data.
```



## References {#References-link}

1. http://www.andrew.cmu.edu/user/achoulde/94842/

