---
title: "Course evaluations and simple linear regression"
author: "Le Nguyen"
date: ""
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

<style type="text/css">
/* Title */
h1.title {
  color: #08738a;
  font-size:40px;
  font-weight: bold;
}
/* Level 1 header */
h1 {
  color: #0ba2c2;
}
/* Level 2 header */
h2 {
  color: #0ba2c2;
}
/* Level 4 header */
h4 {
  font-weight: bold;
}
/* Table of contents */
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #82D0C7;
    border-color: #337ab7;
}
    
</style>

```{r setup, include=FALSE}
# global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA,
                      fig.width=6, 
                      fig.height=5)
```

# Introduction

This is adapted from Lab 6 in Duke's Introduction to Data Science course.

We will analyze what goes into course evaluations and how certain variables
effect the overall score.

To get started, load packages `tidyverse` and `broom`. Install
any packages with code `install.packages("package_name")`.

```{r packages, echo=TRUE}
library(tidyverse)
library(broom)
```


```{r theme-custom, include=FALSE}
theme_custom <- function() {
  theme_bw() +
  theme(text = element_text(color = "#4C6A92"),
        axis.title = element_text(size = 16), 
        title = element_text(size = 20),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.caption = element_text(size = 10))
}
```

# Data

The data were gathered from end of semester student evaluations for a 
large sample of professors from the University of Texas at Austin. 
In addition, six students rated the professors' physical appearance. 
Each row in `evals` contains a different course and the 
columns represent variables about the courses and professors.

Use `read_csv()` to read in the data and save it as an object named `evals`.
The data is available on D2L.


```{r data, echo=TRUE}
evals <- read_csv("evals-mod.csv")
glimpse(evals)
attach(evals)
levels(as.factor(cls_profs))
```

## Data dictionary

**Variable** | **Description**
-------------|---------------------------------------------------------
score |	Average professor evaluation score: (1) very unsatisfactory - (5) excellent
rank |	Rank of professor: teaching, tenure track, tenure
ethnicity |	Ethnicity of professor: not minority, minority
gender |	Gender of professor: female, male
language |	Language of school where professor received education: english or non-english
age |	Age of professor
cls_perc_eval |	Percent of students in class who completed evaluation
cls_did_eval |	Number of students in class who completed evaluation
cls_students |	Total number of students in class
cls_level |	Class level: lower, upper
cls_profs |	Number of professors teaching sections in course in sample: single, multiple
cls_credits	| Number of credits of class: one credit (lab, PE, etc.), multi credit
bty_f1lower |	Beauty rating of professor from lower level female: (1) lowest - (10) highest
bty_f1upper |	Beauty rating of professor from upper level female: (1) lowest - (10) highest
bty_f2upper |	Beauty rating of professor from upper level female: (1) lowest - (10) highest
bty_m1lower |	Beauty rating of professor from lower level male: (1) lowest - (10) highest
bty_m1upper |	Beauty rating of professor from upper level male: (1) lowest - (10) highest
bty_m2upper |	Beauty rating of professor from upper level male: (1) lowest - (10) highest

# Exploratory data analysis

#### Task 1

Create a scatter plot between `score` and a quantitative variable. Comment
on the relationship. What do you think the correlation is between the variables?
```{r}
library(ggplot2)

ggplot(evals, aes(score, age)) + geom_point()

#No correlation, the values look pretty evenly distributed. 
```



#### Task 2

Create a scatter plot between `score` and a categorical variable. Comment
on the relationship. Add `geom_jitter()` to your plot.
What do you notice about the variability in score
between different levels?

```{r}
ggplot(evals, aes(score, gender)) + geom_point() + geom_jitter()

#The score seems equally variable across genders
```


#### Task 3

Create a variable named `avg_bty`. This is the average attractiveness from the
six student scores for each professor. Add this variable to `evals`.
*Hint*: `rowwise()` then `mutate()` then `ungroup()`

```{r}
evals = evals %>%
  rowwise() %>%
  mutate(avg_bty = mean(c_across(bty_f1lower:bty_m2upper)))

```


#### Task 4

Create a scatter plot between `score` and `avg_bty`. Comment
on the relationship. What do you think the correlation is between the variables?
```{r}
ggplot(evals, aes(score, avg_bty)) + geom_point() 

#Slight positive correlation between the two variables.
```

#### Task 5

Modify your plot from Task 4. Add a layer with `geom_jitter()`. Comment
on the relationship.
```{r}
ggplot(evals, aes(score, avg_bty)) + geom_point() + geom_jitter()

#The positive corrolation between the variables is more apparent now.
```
# Model build

#### Task 6

Fit a linear model with function `lm()` between variables `score` and
`avg_bty`. Save it as an object `m.bty`.
What are the estimates for $b_0$ and $b_1$? *Hint*: `tidy()`
```{r}

m.bty <- lm(score ~ avg_bty, data = evals)
m.bty %>%
  broom::tidy()
#b0 = 3.88, b1 = .066
```


#### Task 7

Modify your plot from Task 4 and include the regression line. Pick a different
color than the default of blue.
```{r}
evals %>% 
  ggplot( aes(x = score, y = avg_bty)) + geom_point() + geom_smooth(method=lm, se= FALSE, color = "red")
```


#### Task 8

Interpret the slope of the fitted model. Does the intercept have any practical
meaning within the scope of our data?
```{r}
#The slope of the fit shows the correlation between the variables. In this case there is a small positive correlation between the variables. The intercept shows what average beauty a professor with score zero would have. In this case I think the quantity is meaningless because there isn't a direct corrolation between beauty and score. 
```


#### Task 9

What is the $R^2$ value? Interpret this value. *Hint*: `glance()`
```{r}
m.bty %>%
  glance(r.squared) 

#the r^2 value is .035 which means the fit is terrible, there's a high variability in the regression
```

# Residuals

#### Task 10

Create a plot of the residual value versus `avg_bty`. Comment on what you
observe.

```{r}
m.bty %>%
  augment() %>%
  ggplot(mapping = aes(x =avg_bty,y = .resid)) +geom_point()

#the residuals are fairly uniform, but have a slight trend to them. This means that the type of fit used was correct. 
```

#### Task 11

Compute the sum of squared residuals. Function `augment()` will give you all
the residuals.
```{r}
res = resid(mod.fit)
sum_res = sum(res**2)
sum_res
```



# References {#References-link}

1. http://www2.stat.duke.edu/courses/Spring18/Sta199/labs/lab-06-modelling-course-evals.html

