---
title: "Familarizing with R/RStudio"
author: ""
date: "M1A2 Week 1"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```
# Information

## This database gives information on different demographics for countries in the world.

```{r, echo=FALSE}
u <- "http://www.finley-lab.com/files/data/WorldBank.csv"
WorldBank <- read.csv(u, header = TRUE, stringsAsFactors = FALSE)
head(WorldBank)
```
How many variables are there in the dataset?

```{r, echo=FALSE}
ncol(WorldBank)
```
What are the names of the different variables in the dataset?
```{r, echo=FALSE}
names(WorldBank)

```
```{r, include=FALSE}
#what are the types of the different variables?
str(WorldBank)
```

Identify the types of the different variables.

What is the timeframe of the database?

```{r, echo =FALSE}
data.frame(min(WorldBank$year), max(WorldBank$year))

```

What is the dimension of the dataset?

```{r, echo =FALSE}
dim(WorldBank$year)

```

Extract the information for rows 15 of the database.


Extract the information for rows 55:65 of the database.


##Regions of the World

The major demographic areas included in the database:

```{r}
unique(WorldBank$region)
```

Form a data frame using the latitude and longitude for one unique region. Here the latitube and longitude for *North America* is given below:

```{r, echo=FALSE}
latNA<-unique(WorldBank$latitude[WorldBank$region=="North America"])
longtNA<-unique(WorldBank$longitude[WorldBank$region=="North America"])
data.frame(latNA, longtNA)
```

**Select another region and give the latitude and longitude for that region. Please be sure to mention which region was selected.**


# Analyse *Fertility* and *Life expectancy*

##Visual analysis of fertility and Life-expectancy data for years 1960 and 1975.

**Plots**

**Year 1960**

```{r, echo=FALSE}
# fertility and LifeExpectancy for 1960
fertility_60 <- WorldBank$fertility.rate[WorldBank$year == 1960]

lifeexp_60 <- WorldBank$life.expectancy[WorldBank$year == 1960]
data.frame(fertility_60[1:5],lifeexp_60[1:5])
```


In this section we will plot the data for fertility and life expectancy 

```{r, echo=FALSE}
# fertility and LifeExpectancy for 1960
fertility_60 <- WorldBank$fertility.rate[WorldBank$year == 1960]

lifeexp_60 <- WorldBank$life.expectancy[WorldBank$year == 1960]
```

The simple plot for fertility vs life expectancy for **1960**.

```{r, echo=FALSE}
#plot
plot(lifeexp_60, fertility_60)
```

The enhanced plot of the **1960** data for fertility and life expectancy based on the population size.

```{r, include==FALSE}
#population and region for 1960
pop_60 <- WorldBank$population[WorldBank$year == 1960]
region_60 <- WorldBank$region[WorldBank$year == 1960]
plot(lifeexp_60, fertility_60, type="n")
symbols(lifeexp_60, fertility_60, circles =sqrt(pop_60/pi), inches=0.35,fg=1:8, bg=match(region_60, unique(region_60)))
legend("bottomleft", inset=0.005,
       unique(region_60), fill=1:8, cex=0.8)
```

**Year 1975**

```{r, echo=FALSE}
fertility_75 <- WorldBank$fertility.rate[WorldBank$year == 1975]
lifeexp_75 <- WorldBank$life.expectancy[WorldBank$year == 1975]
data.frame(fertility_75[1:10],lifeexp_75[1:10])
```

The simple plot for fertility vs life expectancy for **1975**.

```{r, echo=FALSE}
plot(lifeexp_75, fertility_75)
```

The enhanced plot of the **1975** data for fertility and life expectancy based on the population size.

```{r, echo=FALSE}
pop_75 <- WorldBank$population[WorldBank$year == 1975]
region_75 <- WorldBank$region[WorldBank$year == 1975]
plot(lifeexp_75, fertility_75, type="n")
symbols(lifeexp_75, fertility_75, circles = sqrt(pop_75/pi), inches=0.35,fg = 1:8,bg=match(region_75, unique(region_75)))
legend("bottomleft", inset=0.005,
       unique(region_75), fill=1:8, cex=0.8)
```

# Comparison of fertility and life-expectancy data for the two years

Comment on the trend, shift and visible change in densities based on population (3-4 lines).

##For years 1965 vs 1970

##For years 1970 vs 2008

