---
title: "Flying with dplyr"
author: ""
date: "M2A4 Week2"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    df_print: paged
---

<style type="text/css">

/* Title */
h1.title {
  color: #6D1DD3;
}

/* Header 1 */
h1 {
  color: #6D1DD3;
}

/* Header 2 */
h2 {
  color: Salmon;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA)
```

# Introduction

We will work with the data frame `flights`, which is included in the 
`nycflights13` package. To get started load `tidyverse` and `nycflights13`
with

```{r packages}
library(tidyverse)
library(nycflights13)
```

You may need to install `nycflights13`. Run `install.packages("nycflights13")`
in your RStudio Console pane.

Package `nycflights13` contains a data frame `flights` that has 
on-time data for all flights that departed NYC (i.e. JFK, LGA or EWR) in 2013.
Take a few minutes to examine the variables and their descriptions with regards
to `flights`. Run `?flights` in your RStudio Console pane.

```{r view}
flights
```

Object `flights` is a tibble. Another way to view the tibble in order to see
all variables is with function `glimpse()`.

```{r glimpse}
glimpse(flights)
```

## Comparison operators

Before you get started, take a few minutes to refresh on some of R's
comparison operators detailed below.

Operator | Description
---------|-------------
`>`  | greater than
`<`  | less than
`>=` | greater than or equal to
`<=` | less than or equal to
`==` | equal to 
`!=` | not equal to
`&`  | and (ex: `(5 > 7) & (6*7 == 42)` will return the value FALSE)
`|`  | or  (ex: `(5 > 7) | (6*7 == 42)` will return the value TRUE)
`%in%` | group membership

To evaluate group membership:

```{r group-member, eval = F}
# Generating the group:
set.seed(634789234)
die.out <- sample(x = 1:6, size = 10, replace = T)
die.out

#Checking for group membership:
die.out %in% c(3, 4)
c(3, 4) %in% die.out

die.out %in% c(1)
c(1) %in% die.out
```

# dplyr

Package dplyr is based on the concept of functions as verbs that 
manipulate data frames.

Function | Action and purpose
---------|-------------------------------
`filter()` | choose rows matching a set of criteria
`slice()` | choose rows using indices
`select()` | choose columns by name
`pull()` | grab a column as a vector
`rename()` | rename specific columns
`arrange()` | reorder rows
`mutate()` | add new variables to the data frame
`transmute()` | create a new data frame with new variables
`distinct()` | filter for unique rows
`sample_n / sample_frac()` | randomly sample rows
`summarise()` | reduce variables to values

## Exercise set 1

Make use of `%>%` operator and any of the functions in package `dplyr` to 
answer the following questions.

1. Filter `flights` for those in January with a destination of
   Detroit Metro (DTW) or Chicago O'Hare (ORD).
```{r 1.1}
flights %>% filter(dest == 'DTW' | dest == 'ORD' & month == 1)
   
```
   
2. Filter `flights` for those before April with a destination that is not
   Detroit Metro (DTW) and had an origin of JFK.
```{r 1.2}
flights %>% filter(dest != 'DTW' & month <= 4 & origin == 'JFK')

```

3. Choose rows 1, 3, 7, 20 from `flights`.
```{r 1.3}
flights %>% slice(c(1,3,7,20))
```

4. Arrange flights by distance and then by departure delay, with the 
   sorting being in descending order in both cases. *Hint:* `desc()`
```{r 4}
flights %>% arrange(desc(flights$distance))
flights %>% arrange(desc(flights$dep_delay))

```
   
5. Select only columns month, origin, and destination from `flights`.
```{r 5}
flights %>% select('month','origin', 'dest')
```

6. Add a new variable to `flights` called `gain`, where `gain` is the
   arrival delay minus the departure delay.
```{r 6}
gain = flights$arr_delay - flights$dep_delay
cbind(flights, gain)
```

   
7. Use summarise to obtain the mean departure delay and mean arrival delay for all flights with an origin of EWR.
```{r 7}
flights_EWR = subset(flights, origin == 'EWR')
flights_EWR
flights_EWR %>% summarise(dep_delay)
flights_EWR %>% summarise(arr_delay)

```

   
## Exercise set 2

Grouping adds substantially to the power of the `dplyr` functions. We will 
focus on using `summarise()` with `group_by()`, but grouping also can be 
used with other `dplyr` functions.

1. Create a data frame which contains the number of flights and the mean arrival delay for flights on carrier UA (United Airlines) whose destination is O'Hare Airport (ORD). The number of flights and mean arrival delay is calculated separately for flights out of each of the origin airports.
```{r 2.1}
flights_sorted = subset(flights, dest == 'ORD' & carrier == 'UA')
flights_sorted

length(flights_sorted$arr_delay)
mean(flights_sorted$arr_delay)
```
```{r}

```
   


# References

1. https://cran.r-project.org/web/packages/dplyr/vignettes/introduction.html


